---
title: "Soccer Analysis Using Statsbomb Data"
date: "04-20-2026"
output: 
  html_document: 
    theme: cerulean
---
## **Introduction**

The data is sourced from StatsBomb's Open Data repository via their R package. It contains highly detailed, event level data for various soccer competitions. For this analysis, we are specifically filtering the data to look at UEFA Champions League matches.

* A link to the data repository [StatsBomb Open Data on GitHub](https://github.com/statsbomb/open-data)

## **Questions**

* To understand tactical trends, team performance, and player behaviors within the Champions League.

  * *Which players were caught in an offside position the most frequently?*
  * *?*
  * *?*
  * *?*

## **Key Terms for Unfiltered Data & Filtered Data**

* `player.name`: Name of the primary player involved in the event

* `type.name`: The primary category of the event (e.g., "Pass", "Offside", "Shot")

* `pass.outcome.name`: The result of a pass attempt (e.g., "Pass Offside")

* `pass.recipient.name`: The intended target of a pass (used to identify who was offside during a pass)

* `offside_player`: A calculated field identifying the specific player caught offside

* `offside_count`: The total number of times a specific player was caught offside

## **Key Values and Calculations:**




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r 1}
#install.packages("devtools")
#install.packages("remotes")
#devtools::install_github("statsbomb/StatsBombR")

library(tidyverse)
library(dplyr)

# Required packages by StatsBombR
library(stringi)
library(rvest)
library(RCurl)
library(foreach)
library(iterators)
library(parallel)
library(httr)
library(jsonlite)
library(sp)

library(StatsBombR)
```


```{r}
# Load all competitions
Comp <- FreeCompetitions()
glimpse(Comp)

# Show all competition names 
unique(Comp$competition_name)
```

```{r}
champLeagues_Comp <- Comp %>% 
  filter(competition_name == "Champions League")

Matches <- FreeMatches(champLeagues_Comp)


champLeagues_events <- list()

for (i in 1:nrow(Matches)) {
  match_data <- get.matchFree(Matches[i,])
  champLeagues_events[[i]] <- match_data  
  print(paste("Loaded match", i, "of", nrow(Matches)))
}
champLeagues_events <- bind_rows(champLeagues_events)

```
## **Most offside positions in the Champions League**
```{r 3}
offside_players <- champLeagues_events %>%
  filter(pass.outcome.name == "Pass Offside" | type.name == "Offside") %>%
  mutate(offside_player = ifelse(type.name == "Pass", pass.recipient.name, player.name)) %>%
  filter(!is.na(offside_player)) %>%
  group_by(offside_player) %>%
  summarise(offside_count = n(), .groups = 'drop') %>%
  arrange(desc(offside_count))

head(offside_players, 10)

top_10_offsides <- head(offside_players, 10)

ggplot(top_10_offsides, aes(x = reorder(offside_player, offside_count), y = offside_count)) +
  geom_bar(stat = "identity", fill = "#0E1E5B", width = 0.7) +
  coord_flip() +
  labs(
    title = "Most Offsides in the Champions League",
    subtitle = "Players caught in an offside position",
    x = "Player",
    y = "Number of Offsides"
  ) +
  theme_minimal()
```
## **Average Goals per Match: Home vs. Away**
```{r 4}
Matches <- Matches %>%  
  mutate(total_goals = home_score + away_score)

avg_total_goals <- mean(Matches$total_goals, na.rm = T)
avg_home_goals <- mean(Matches$home_score, na.rm = T)
avg_away_goals <- mean(Matches$away_score, na.rm = T)

print(paste("Average total goals per match:", round(avg_total_goals, 2)))
print(paste("Average home team goals:", round(avg_home_goals, 2)))
print(paste("Average away team goals:", round(avg_away_goals, 2)))

ggplot(Matches, aes(x = home_score, y = away_score)) +
  geom_jitter(width = 0.2, height = 0.2, alpha = 0.6, color = "blue", size = 3) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  
  geom_abline(intercept = 0, slope = 0.5, linetype = "dashed", color = "green", size = 1) +
  geom_abline(intercept = 0, slope = 2, linetype = "dashed", color = "red", size = 1) +
  
  labs(
    title = "Match Results: Home Goals vs. Away Goals",
    subtitle = "Black line = Draw | Green line = Slope 0.5 | Red line = Slope 2",
    x = "Home Team Goals",
    y = "Away Team Goals"
  ) +
  
  coord_fixed(ratio = 1) +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"))
```


### **Does Defense Win More Games?**

#To determine if defense correlates with winning, we can measure the volume of defensive actions a team performs in a match (Tackles, Interceptions, Clearances, Blocks, and Pressures) and compare it to the match outcome.

```{r defense-analysis, warning=FALSE, message=FALSE}
defensive_events <- champLeagues_events %>%
  filter(type.name %in% c("Pressure", "Tackle", "Interception", "Clearance", "Block")) %>%
  group_by(match_id, team.name) %>%
  summarise(defensive_actions = n(), .groups = 'drop')

team_match_outcomes <- bind_rows(
  Matches %>% select(match_id, team.name = home_team.home_team_name,
                     goals_for = home_score, goals_against = away_score),
  Matches %>% select(match_id, team.name = away_team.away_team_name,
                     goals_for = away_score, goals_against = home_score)
) %>% mutate(outcome = case_when(
    goals_for > goals_against ~ "Win",
    goals_for < goals_against ~ "Loss",
    T ~ "Draw"
  ))

team_match_outcomes %>%
  inner_join(defensive_events, by = c("match_id", "team.name")) %>%
  group_by(outcome) %>%
  summarise(avg_actions = mean(defensive_actions, na.rm = T), .groups = 'drop') %>%
  mutate(outcome = factor(outcome, levels = c("Loss", "Draw", "Win"))) %>%
  ggplot(aes(x = outcome, y = avg_actions, fill = outcome)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(avg_actions, 1)), vjust = -0.5, size = 3, fontface = "bold") +
  scale_fill_manual(values = c("Win" = "green", "Draw" = "blue", "Loss" = "red")) +
  labs(
    title = "Average Defensive Actions by Match Outcome",
    subtitle = "Tackles, Interceptions, Clearances, Blocks, and Pressures combined",
    x = "Match Outcome", y = "Average Defensive Actions"
  ) +
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(size = 14, face = "bold"))
```




